\systemnamenospace: Head Orientation Based Selection\\in Physical Spaces

### 2014-06-28 doc/sui14/applications.tex
use 0.95 for all figures
shorten the paper to fit 8 pages

We implemented a prototype of this remote controller for three physical devices: a lamp and fan that could be switched on or off and a smart TV control with playback, volume and navigation controls (see Figure~\ref{fig:smart-home}). We switched discrete appliances using relays, and implemented controls for a smart TV with a laptop connected to a 30\inch{} display. User interfaces for each were pre-defined in our application.

@@

\begin{figure}[t]
\centering
\includegraphics[width=1\columnwidth]{figures/smarthome-scenario.pdf}
\caption{In the smart home scenario, we have built three smart appliances for a user experience study. The interface supports both simple on/off appliances (lamp or fan) and multi-functional appliances (TV, for example).}
\label{fig:smart-home}
\end{figure}

@@

\changes{Though our solution successfully solves the problem of {\em target selection}, we have not fully addressed the issue of controlling smart devices. One user reported that \studyquote{for more complex appliances (TV), interaction was more difficult}. A separate work on interaction techniques for head-worn computing devices is needed to address these issues, while we keep our focus on target selection. When the device is more complicated, it is possible that the best solution may be a hybrid-- to select the appliance with HOBS and control it with a phone.} \claire{ Is it possible to say that the number of complicated devices that you want to remote control is limited almost exclusively to the tv? I can't think of too many of these devices and our system works very well with more simple devices, which seem to be more common.} \sean{ I wouldn't state that. We could be challenged if the reviewers bring up smart fridge, microwave, etc.}

>>>>

Our prototype includes three smart devices: a lamp and a fan that could be switched on and off; a smart TV with playback, volume and navigation controls (see Figure~\ref{fig:smart-home}). The on/off appliances are using relays, and the smart TV is a 30\inch{} display connected to a laptop. User interfaces for each were pre-defined in our application.

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.95\columnwidth]{figures/smarthome-scenario.pdf}
\caption{In the smart home scenario, we have built three smart appliances for a user experience study. The interface supports both simple on/off appliances (lamp or fan) and multi-functional appliances (TV, for example).}
\label{fig:smart-home}
\end{figure}

@@

\changes{Though our solution successfully solves the problem of {\em target selection}, we have not fully addressed the issue of controlling smart devices. One user reported that \studyquote{for more complex appliances (TV), interaction was more difficult}. A separate work on interaction techniques for head-worn computing devices is needed to address these issues, while we keep our focus on target selection.} \ben{ Suggest to remove this one: When the device is more complicated, it is possible that a better solution may be a hybrid -- to select the appliance with HOBS and control it with a phone.} \claire{ Is it possible to say that the number of complicated devices that you want to remote control is limited almost exclusively to the tv? I can't think of too many of these devices and our system works very well with more simple devices, which seem to be more common.} \sean{ I wouldn't state that. We could be challenged if the reviewers bring up smart fridge, microwave, etc.}

### 2014-06-28 doc/sui14/discussion.tex
edits
refining applications (with limitation for complex devices) and commented 'comments' in disucssion
shorten the paper to fit 8 pages

%!TEX root = uist14.tex
\section{Discussion}
\label{sec:discussion}
%%%NOTES FROM YANG%%%%
The rapid development of sensing technologies has created many opportunities for new ways to interact with smart objects. In our exploration of the design space of head orientation-based target selection, we carefully selected sensing techniques that are readily available and easy to deploy; and we added complexity to our system only when necessary.

@@

A primary goal of this paper was to provide an effective and efficient method for target selections in physical spaces. Targeting is a fundamental building block across many interaction tasks - it has a significant impact on user experiences collectively and can provide seamless interaction when designed well.

We formalized a scan and refine model of head orientation-based selection (Equation~\ref{eq:time}). We first introduced head orientation as an alternative to list selection and showed that {\em scanning} can outperform list selection. Our redesigns then focused on the case where refinement is needed. The two ways to reduce refinement time are 1) to reduce $P(refine)$, the probability that a manual refinement is necessary; and 2) to reduce $t_{refine}$, the time required to perform the refinement interaction itself. Using IR intensity readings addresses both these terms, as it can be used to both avoid showing refinement dialogs, and to optimize their display when they are needed.

@@

\subsection{Limitations}
We introduced new refinement operations based on IR intensity and spatial adjacency tracking that improve the performance of head orientation in environments with dense targets. These techniques have several limitations.

First, IR intensity measurements only work within the dynamic range of our sensor. Additional strong IR sources like direct sunlight may saturate the sensor and make discrimination impossible. Second, our adjacency map is built assuming stable relative target locations. If targets move, the map will have to be recalculated. This may be done incrementally during everyday interactions, but we have not yet tackled this challenge. We also have not yet evaluated how robust the adjacency map is when the user changes her location or view angle.

Finally, we acknowledge several limitations of our study design: we have not yet systematically studied target density variation, our study was performed in a lab environment, and only measured first use. We have also not investigated error rates in detail. Future work should study how the technique applies in realistic settings over longer periods of time.
%Point 3: Limitation
%Highlight the novelty of technique 2 & 3 while point out its weakness:
%- How sunlight will impact the techniques
%- Construct the adjacency mapping, what if objects move
%- How robust its adjacency mapping is when the user changes her location or view angle.
%- Outline programming by demonstration opportunities for future research
%- Study limitation: density variation, realistic settings, longitudinal studies to understand how users choose between our techniques and traditional list-based solution.

>>>>

%!TEX root = uist14.tex
\section{Discussion}
\label{sec:discussion}
%%%NOTES FROM YANG%%%%
The rapid development of sensing technologies has created many opportunities for new ways to interact with smart objects. In our exploration of the design space of HOBS, we carefully selected sensing techniques that are readily available and easy to deploy; and we added complexity to our system only when necessary.

@@

A primary goal of this paper was to provide an effective and efficient method for target selections in physical spaces. Targeting is a fundamental building block across many interaction tasks -- it has a significant impact on user experiences collectively and can provide seamless interaction when designed well.

We formalized a {\em scan} and {\em refine} model of head orientation-based selection (Equation~\ref{eq:time}). We first introduced head orientation as an alternative to list selection and showed that {\em scanning} can outperform list selection. Our redesigns then focused on the case where refinement is needed. The two ways to reduce refinement time are 1) to reduce $P(refine)$, the probability that a manual refinement is necessary; and 2) to reduce $t_{refine}$, the time required to perform the refinement interaction itself. Using IR intensity readings addresses both these terms, as it can be used to both avoid showing refinement dialogs, and to optimize their display when they are needed.

@@

\subsection{Limitations}

Our system faces a few limitations. First, IR intensity measurements only work within the dynamic range of our sensor. Additional strong IR sources like direct sunlight may saturate the sensor and make discrimination impossible. Second, our adjacency map is built assuming stable relative target locations. If targets move, the map will have to be recalculated. This may be done incrementally during everyday interactions, but we have not yet tackled this challenge.

We also acknowledge several limitations of our study design: we have not yet systematically studied target density variation, our study was performed in a lab environment, and only measured first use. Future work should study how the technique applies in realistic settings over longer periods of time.

%Point 3: Limitation
%Highlight the novelty of technique 2 & 3 while point out its weakness:
%- How sunlight will impact the techniques
%- Construct the adjacency mapping, what if objects move
%- How robust its adjacency mapping is when the user changes her location or view angle.
%- Outline programming by demonstration opportunities for future research
%- Study limitation: density variation, realistic settings, longitudinal studies to understand how users choose between our techniques and traditional list-based solution.

%% \bjoern{removed CV subsection; handled in related work}

%\subsection{Computer-vision Solution?}
%Reviewers suggest considering computer vision-based techniques with markers in the environment (such as QR codes, or Bokode [Mohan] which extends the working distance to a few meters for visual tags). We considered this direction and found several drawbacks, including computational complexity, low accuracy when targets are far, the absence of real-time environment feedback with passive tags, and aesthetic issues with QR codes. While we don't claim our choice of IR is optimal, it has important advantages -- it is low-cost, readily available and quite suitable for area selection. We can weaken our claim and add explanations of how disambiguation techniques might generalize to other methods of implementing target identification.

### 2014-06-28 doc/sui14/relatedwork.tex
use 0.95 for all figures
shorten the paper to fit 8 pages

%!TEX root = sui14.tex
\section{Background and Related Work}
Our approach is related to head- and eye-controlled interfaces, area cursors and prior work on hardware devices for pointing in physical spaces.

\vfill
\ben{make sure no section title appears at the end of any column.}
\subsection{Head and Gaze Input}
\changes{Head movement has long been used for virtual camera control in VR applications~\cite{pausch_user_1993} and as an assistive input technology for cursor control of desktop applications~\cite{radwin1990method}. However, it is notable that human neck muscles have a lower bandwidth than other muscle groups, e.g., the wrist~\cite{card_morphological_1991}.}
%
%Card, summarizing other work, shows that neck muscles are a poor muscle group for pointing in general: neck muscles only have a bandwidth of about $4.2bits/s$ (compared to $23bits/s$ of wrist muscles used by a standard mouse~\cite{Card:1991:MAD:123078.128726}. 
%
Prior work often focused on head orientation for controlling graphical interfaces; in contrast, we apply this modality to selection in physical spaces.

Gaze can also be used to control graphical user interfaces~\cite{kumar2007eyepoint}. While there are wearable gaze trackers~\cite{bulling2009wearable}, turning information about a concrete point in space where a user is looking into a selection requires a map with known target locations.
Our system works through point-to-point IR communication and does not require an {\em a priori} map or markers. Target objects in the environment can also be equipped with individual cameras that watch the user~\cite{smith2013gaze,vertegaal2005media}. Such an approach can enable similar benefits as our approach, but is computationally more expensive than our single-pixel sensor solution and may not work at greater distances or angles, because it relies on finding the user's pupils in a camera image.

@@

\subsection{Area Cursors}
\changes{
In 2D area cursors for GUIs, the activation area of the cursor is enlarged, which facilitates acquiring smaller targets~\cite{kabbash1995prince}. We argue that head orientation pointing has analogous characteristics (limited pointing performance and accuracy). Area cursors are especially appropriate for individuals with motor control impairments or difficulties~\cite{worden1997making,findlater2010enhanced}. Similar ideas have also been extended into 3D to provide selection with progressive refinement in 3D scenes~
\cite{bacim2013design}. In all area and space cursors, the large activation area can lead to multiple targets being selected and disambiguation is needed. This paper describes the trade-offs between several disambiguation approaches.

@@

\subsection{Pointing in Physical Spaces}
Rukzio ~\cite{rukzio_experimental_2006} studied alternative methods for selecting devices in physical spaces and found that users strongly preferred either tapping target appliances with a mobile device or pointing at a distance to browsing a list.

Several approaches to spatial selection with handheld devices~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012,kemp_point-and-click_2008} or finger-worn devices~\cite{merrill_augmenting_2007} exist. 
%and exchanging information with smart infrastructure sensor networks ~\cite{lifton_tricorder:_2007,mittal_ubicorder:_2011,costanza_sensortune:_2010}. 
In some techniques, users select objects of interest with laser pointers. The laser dot provides immediate visual feedback to the user about what is being selected; however, its small target area makes it poorly matched to head orientation input.
Other approaches rely on virtual room models in which a user's location is estimated using IMU-based orientation sensing~\cite{wilson_xwand:_2003,lifton_tricorder:_2007} -- in contrast, our technique does not require a static map ahead of time.
%Other targeting systems use IR with handheld pointers~\cite{swindells_that_2002} as well as wearable devices such as rings and Bluetooth audio earpieces~\cite{merrill_augmenting_2007} to connect to smart devices. 
Our system tackles an unresolved issue of prior approaches -- navigating an area dense with potential targets and resolving selection ambiguity.

>>>>

%!TEX root = sui14.tex
\vfill
\ben{make sure no section title appears at the end of any column.}

\section{Background and Related Work}
Our approach is related to head- and eye-controlled interfaces, area cursors, pointing in physical spaces and computer vision-based selections.

\subsection{Head and Gaze Input}
\changes{Head movement has long been used for virtual camera control in VR applications~\cite{pausch_user_1993} and as an assistive input technology for cursor control of desktop applications~\cite{radwin1990method}. However, it is notable that human neck muscles have a lower bandwidth than other muscle groups, e.g., the wrist~\cite{card_morphological_1991}.}
%
%Card, summarizing other work, shows that neck muscles are a poor muscle group for pointing in general: neck muscles only have a bandwidth of about $4.2bits/s$ (compared to $23bits/s$ of wrist muscles used by a standard mouse~\cite{Card:1991:MAD:123078.128726}. 
%
Prior work often focused on head orientation for controlling graphical interfaces; in contrast, we apply this modality to selection in physical spaces.

Gaze can also be used to control graphical user interfaces~\cite{kumar2007eyepoint}. While there are wearable gaze trackers~\cite{bulling2009wearable}, turning information about a concrete point in space where a user is looking into a selection requires a map with known target locations.
Our system works through point-to-point IR communication and does not require an {\em a priori} map or markers. Target objects in the environment can also be equipped with individual cameras that watch the user~\cite{smith2013gaze,vertegaal2005media}. Such an approach can enable similar benefits as our approach, but is computationally more expensive and may not work at greater distances or angles, because it relies on finding the user's pupils in a camera image.

@@

\subsection{Area Cursors}
\changes{
In 2D area cursors for GUIs, the activation area of the cursor is enlarged, which facilitates acquiring smaller targets~\cite{kabbash1995prince}. We argue that head orientation pointing has analogous characteristics (limited pointing performance and accuracy). Area cursors are especially appropriate for individuals with motor control impairments or difficulties~\cite{worden1997making,findlater2010enhanced}. Similar ideas have also been extended into 3D to provide selection with progressive refinement in 3D scenes~\cite{bacim2013design}. All area and space cursors necessitate disambiguation when there are multiple targets and no clear winner. This paper describes the trade-offs between several disambiguation approaches.

@@

\subsection{Pointing in Physical Spaces}
Rukzio ~\cite{rukzio_experimental_2006} studied alternative methods for selecting devices in physical spaces and found that users strongly preferred either tapping target appliances with a mobile device or pointing at a distance to browsing a list. Several other approaches to spatial selection with handheld devices~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012,kemp_point-and-click_2008} or finger-worn devices~\cite{merrill_augmenting_2007} exist. 
%and exchanging information with smart infrastructure sensor networks ~\cite{lifton_tricorder:_2007,mittal_ubicorder:_2011,costanza_sensortune:_2010}. 
In some techniques, users select objects of interest with laser pointers; however, the laser dot's small target area makes it poorly matched to head orientation input.
Other approaches rely on virtual room models in which a user's location is estimated using IMU-based orientation sensing~\cite{wilson_xwand:_2003,lifton_tricorder:_2007} -- in contrast, our technique does not require a static map ahead of time.
%Other targeting systems use IR with handheld pointers~\cite{swindells_that_2002} as well as wearable devices such as rings and Bluetooth audio earpieces~\cite{merrill_augmenting_2007} to connect to smart devices. 
Our system tackles an unresolved issue of prior approaches -- navigating an area dense with potential targets and resolving selection ambiguity.
